---
title: "Transform-Assert"
author: "John Yang"
output: html_notebook
---

# First install the packages
```{r}
install.packages("tidyverse")
install.packages("nycflights13")
```
# Now call the libraries

````{r}
library(nycflights13)
library(tidyverse)
library(tidyr)
```
# Relational Data: look at all these tables

````{r}
flights 
airlines 
airports 
planes 
weather 
```
Individual excercise from last time: Join the flights and weather tables using c("year", "month", "day", "hour", "origin")

````{r}

flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)

flights2

weather

weather %>%
count(year, month, day, hour, origin) %>% 
  filter(n > 1) 

flights2 %>% 
  left_join(weather, by = c("year", "month", "day", "hour", "origin"))

# Question for you: why the total number of observations remian the same after left_join?

```


What you have already known: You can define a group of variables as a key

````{r}
flights2
weather

flights2 %>% 
  left_join(weather, by = c("year", "month", "day", "hour", "origin"))
```

A named character vector: by = c("a" = "b"). This will match variable a in table x to variable b in table y.

````{r}
flights2
airports

flights2 %>% 
  left_join(airports, c("dest" = "faa"))

flights2 %>% 
  left_join(airports, c("origin" = "faa"))

```

# Filter joins: semi-join and anti-join

````{r}
top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)

top_dest

flights 

flights %>% 
  semi_join(top_dest)

flights %>% 
  anti_join(top_dest)
```

Individual excercise: 

Using a group of variables as a key (year, month, day, hour, origin), find the observations in flights2 tibble where no weather information is available.

````{r}

flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)

# Your code goes here.
flights2

weather

all_weather <- weather %>%
count(year, month, day, hour, origin) 


flights2 %>% 
  anti_join(all_weather)
 


```

