---
title: "vectors"
output: html_document
---




## Problem 1: Atomic Vectors and Vectorized Functions (20 pts)

The `rivers` data is a numeric vector containing the lengths in miles of 141 rivers in North America. 

```{r}
data("rivers")
```

a) What is a vectorized function and 2) why are they so useful? 

Vectorized function can operate on all elements of the whole vector at a time, and it doesn't need to loop through and act on each element one at a time.
 
Because it makes code more concise, easy to read, and less error prone. The multiplication happened to each element of the vector.

b) Write a function that will convert the river length from miles to km using the conversion 1 mile = 1.60934 km. The function should take in a vector and return a new vector. Note: Do not write a loop in this function. 

```{r}
rivers_in_km <- function(x){
  k = 1.609344
  rivers_km = rivers*k
  rivers_km
}
rivers_in_km(rivers)
```



## Problem 2: More vectors (20 pts)

a) Write a function that will take a vector and return the number of elements with values greater than 1,000. For example, if the input vector is c(108, 2060, 3442, 1000), the function would return 2. 

In the body of your function, be sure to use vector operations. You don't need an if-else statement or a loop! (See section 20.3.1 of R for Data Science)

```{r}
greater1000 <- function(x){
  sum(x>1000)
}
```


b) Test your function with the `rivers` data to answer: How many rivers are longer than 1,000 miles?

```{r}
greater1000(rivers)
```




## Problem 3: Factors (10 pts)

Factors are a type of vector known as an augmented vector. Create a factor called `river_level` that categorizes river lengths into the following levels:

* short: < 500 miles long
* medium: >= 500 miles and < 1000 miles
* large: >= 1000 miles

Hint: look up the cut() function. 

```{r}
river_level<-cut(rivers,breaks=c(0,500,1000,Inf),
                 labels=c("short","medium","large"), 
                 include.lowest = TRUE,right = FALSE)
river_level
```

## Problem 4: Indexing with Lists (25 pts)

Below is a list called `tweet_list`, which contains information about the top 20 most retweeted tweets on the social media platform Twitter along with some meta data about the list. To receive credit, you must use `tweet_list` in the expressions for parts a) and b) below. 

```{r}
# Number of retweets per tweet according to Twitter's rounding
num_rts <- c(1140000, 1500000, 1030000, 1600000, 1300000, 1400000, 3400000, 2600000, 3100000, 1020000, 1160000, 1010000, 938600, 1800000, 1040000, 4400000, 3800000, 999900, 1030000, 912000)
# Author of the tweet
tweet_author <- c('MrBeastYT', 'Rubiu5', 'onedirection', 'BarackObama', 'BTS_twt', 'BTS_twt', 'carterjwm', 'Louis_Tomlinson', 'TheEllenShow', 'ArianaGrande', 'BTS_twt', 'RandallTime', 'DennyJA_WORLD', 'Rubiu5', 'PennStateIFC', 'yousuck2020', 'yousuck2020', 'RyanSesselman', 'BTS_twt', 'BTS_twt')
# List
tweet_list <- list(c("Top 20 Most Retweeted", "July 30, 2020"), num_rts, tweet_author)
```

a) Write an expression to find the sum of `num_rts` across all tweets.

```{r}
sum(tweet_list[[2]])
```


b) It has been discovered that twitter user MrBeastYT bought retweets, artificially inflating the number of retweets his tweet received. Subtract 500,000 purchased retweets from his retweet count (`num_rts`).

```{r}
tweet_list[[2]][1] <- tweet_list[[2]][1] - 500000
tweet_list
```