---
title: "Join"
output: html_document
---

```{r}
library(readr)
library(dplyr)
```

## 1, Right join and left join table_2 to table_1. What's the difference between these two joined table? 

```{r}
df1 <- read.delim("table_1.csv")
df2 <- read.delim("table_2.csv")

df1 %>% 
right_join(df2, by = "state")

df1 %>% 
left_join(df2, by = "state")
```

We can see the difference that the `right_join` keeps all the observation of `table_2`, and combines the same observations with the variable "population" and "ev". And then for the observations which `table_1` doesn't have, the "NA" would show right there in "population". The same, `left_join` keeps all the observation of `table_1`, and also combines the same observations with 2 variables. For the observations which `table_2` doesn't have, the "NA" would show in "ev".


## 2, Use appropriate join to get the intersection table of table_1 and table_2.

```{r}
inner_join(df1,df2)
```


## 3, Use appropriate join to get the union table of table_1 and table_2.

```{r}
full_join(df1,df2)
```


## 4, Use appropriate join to keep the part of first table for which we have information in the second.

```{r}
df1 %>% 
semi_join(df2, by= "state")
```


## 5, Use appropriate join to achieve opposite result from question 4.

```{r}
df1 %>% 
anti_join(df2, by= "state")
```


