---
title: "Transform-Assert"
author: "John Yang"
output: html_notebook
---

# First install the packages
```{r}
install.packages("tidyverse")
install.packages("nycflights13")
```

# Now call the libraries

````{r}
library(nycflights13)
library(tidyverse)
```

Using pipe (%>%) to simplify your code:

Now, suppose we are trying to generate a new tibble that contains all flights in July and adds a new column, reader = "John".   

````{r}
flights 
(flights_July = filter(flights, month == 7))
(flights_July_ZY = mutate(flights_July, reader="John"))
```

Let's use Pipe to simplify the code with three steps:
1. remove the name of each newly generated tibbles
2. remove the first argument of each operation verbs
3. add "%>%"

````{r}
flights %>% 
  filter(month == 7) %>%
  mutate(reader="John")
```

Group Challenge: Investigate the relationship between the distance and average delay for each location in July？
Hint: Use summarise, group_by, filter, arrange

````{r}
flights %>% 
  filter(month == 7) %>% 
  group_by(dest) %>% 
  summarise(dist = mean(distance, na.rm= TRUE), delay = mean(arr_delay, na.rm = TRUE)) %>% 
  arrange(desc(delay))
  
```

# Apply your skills to address covid19 statistics

````{r}
install.packages("covid19.analytics")
```

````{r}
library(covid19.analytics)
library(tidyverse)
```

````{r}
# obtain time series data for "confirmed" cases
(covid19.confirmed.cases <- covid19.data("ts-confirmed"))
```

````{r}
# First check the number of variables. A function that helps us explore the number of days/weeks between two dates. 
difftime("2021-9-6", "2020-1-22", units = "days")
```

Individual excercise (3 mins)

````{r}
# Find the top ten countries that have the largest number of confirmed cases on 2021-09-02 (recommend using pipe!)
# Hint-1: the following verbs are recommended: select, group_by, summarise, filter
# Hint-2: filter(rank( desc(X) )) <= 10 will help you identify the top ten values of X

covid19.confirmed.cases %>% 
  select(Country.Region, `2021-09-02`) %>% 
  group_by(Country.Region) %>% 
  summarise(case_sum_02 = sum(`2021-09-02`, na.rm = TRUE)) %>% 
  filter(rank(desc(case_sum_02)) <= 10 )
    
```

Individual excercise (5 mins): how to define daily growth rate?

````{r}
# Adapt the above code to find the top ten countries that have the highest daily growth rate of confirmed cases on 2021-09-02. 
#Hint-1: define a new variable, growth_rate, either within summarise() or using mutate()
#Hint-2: daily growth rate = (cases_day_2 - cases_day_1)/cases_day_2

covid19.confirmed.cases %>% 
  select(Country.Region, `2021-09-02`,`2021-09-01` ) %>% 
  group_by(Country.Region) %>% 
  summarise(case_sum_02 = sum(`2021-09-02`, na.rm = TRUE), case_sum_01 = sum(`2021-09-01`, na.rm = TRUE)) %>%
  mutate(growth_rate = (case_sum_02 - case_sum_01) / case_sum_01 ) %>% 
  filter(rank(desc(growth_rate)) <= 10 )

```

# Now let's take a look at the US data!

````{r}
(CV_US = covid19.US.data() )
```

Use across() function to select a range of variables

````{r}
# Operation by row
# Create a new Table, CV_US_Aug, that contains the by-state sum of daily confirmed cases from 2021-08-01 to 2021-08-31


(CV_US_Aug <- CV_US %>% 
    group_by(Province_State) %>% 
    summarise(across(`2021-08-01`:`2021-08-31`, sum))
)

```

Individual excercise (3 mins): use rowSums() to sum by row.

````{r}
# Find the monthly sum of confirmed cases for each US state in August, 2021. 
# Hint-1: rowSums(across(`date-1`:`date-2`), na.rm = TRUE) will return the by-row sum of values in a tibble. 

CV_US_Aug %>% 
    group_by(Province_State) %>%
    summarise( Sum_Aug_2021 = rowSums(across(`2021-08-01`:`2021-08-31`, na.rm = TRUE)))
# Your code goes here
```

Group challenge*: Identify the top ten states that have the highest mean growth rate of COVID cases in August, 2021.  

Hint-1: across() allows arithmetic operations for a range of variables. 
Hint-2: rowMeans() will return the by-row average of values in a tibble

````{r}
CV_US %>% 
   group_by(Province_State) %>%
   summarise( Sum_Aug_2021 = rowSums(across(`2021-08-01`:`2021-08-31`, na.rm = TRUE))
              )
   mutate( -(across`2021-07-31`:`2021-08-30`, sum)/(across`2021-07-31`:`2021-08-30`, sum))
```


