---
title: "assignment-3"
output: github_document
---
```{r}
library(tidyverse)
library(readr) 
library(ggplot2)
library(dplyr)
```

```{r}
data<- read.csv('train.csv')
view(data)
```


# 1

Find out the survive vs death ratio with respect to gender.And what's the age distribution for each gender.

```{r}
data %>% 
  group_by(Sex) %>% 
  summarise(sur_vs_dea_ratio= sum(Survived, na.rm = TRUE)/(n() - sum(Survived,na.rm = TRUE)))
data %>% 
  na.omit %>% 
  count(Sex, Age)
f_p<-data %>% 
  na.omit %>% 
  filter(Sex=="female") 
  hist(f_p$Age, xlab = "age", main = "Age Distribution of Female")
  ggplot(f_p, aes(x= Age)) + geom_histogram(bins = 30) + labs(title = "Age Distribution of Female") 
m_p<-data %>% 
  na.omit %>% 
  filter(Sex=="male") 
  hist(m_p$Age, xlab = "age", main = "Age Distribution of Male")
  ggplot(m_p, aes(x= Age)) + geom_histogram(bins = 30) + labs(title = "Age Distribution of Male") 

```


# 2

What is the mean age with respect to survived.

```{r}
data %>% 
  filter(Survived == 1) %>% 
  group_by(Survived) %>% 
  summarise(mean_age = mean(Age,na.rm = TRUE))
```


# 3

hint: count, sort, arrange.

Find the 75% percentile of age with respect to survived column, then add a new column called age_bracket. We put major in the age_bracket if's the age is in the 75% percentile and minor for the rest 25% percentile. 

```{r}
data %>% 
  na.omit %>%
  filter(Survived == 1) %>% 
  count(Survived, Age) %>% 
  arrange(desc(n)) %>% 
  mutate(cumsum_n = cumsum(`n`)) %>% 
  mutate(age_bracket = ifelse(cumsum_n >= 0.75*sum(n),"minor","major"))

data %>% 
  na.omit %>%
  filter(Survived == 0) %>% 
  count(Survived, Age) %>% 
  arrange(desc(n)) %>% 
  mutate(cumsum_n = cumsum(`n`)) %>% 
  mutate(age_bracket = ifelse(cumsum_n >= 0.75*sum(n),"minor","major"))

```

75% percentile of age in (Survived == 1) is 45 years old, and 75% percentile of age in (Survived == 0) is 38 years old.
