---
title: "Transform-Assert"
author: "John Yang"
output: html_notebook
---

# First install the packages
```{r}
install.packages("tidyverse")
install.packages("nycflights13")
```
# Now call the libraries

````{r}
library(nycflights13)
library(tidyverse)
library(tidyr)
```

# Relational Data: look at all these tables

````{r}
flights 
airlines 
airports 
planes 
weather 
```
# Finding a Key!

# Let's try year, month, day, and flight

````{r}
flights %>% 
  count(year, month, day, flight) %>% 
  filter(n > 1)
```

# Let's try year, month, day, tailnum

````{r}
flights %>% 
  count(year, month, day,tailnum) %>% 
  filter(n > 1)
```

Individual challenge (5 mins): Add a surrogate key, index, to flights
Hint-1: index should reflect the index of the observation like (1,2,...,336776)
Hint-2: c(1:N) will return a vector of integers containing (1,2,...,N)

````{r}
xnxn <- flights %>% 
      mutate(index=c(1:n()))


         
```
Quick show of an example for "join".

````{r}
(flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier))

# Task: Add the full airline name to the flights2 data
flights2 %>%
  select(-origin, -dest) %>% 
  left_join(airlines, by = "carrier")
```

Individual excercise: jointhe flights and weather tables using c("year", "month", "day", "hour", "origin")

````{r}
weather
flights2 %>% 
  left_join(weather, by = c("year", "month", "day", "hour", "origin"))
```




