---
title: "Transform-Assert"
author: "John Yang"
output: html_notebook
---

# First install the packages
```{r}
install.packages("tidyverse")
install.packages("nycflights13")
```
# Now call the libraries

````{r}
library(nycflights13)
library(tidyverse)
library(tidyr)
```

Key properties of a vector

````{r}
# Use typeof() to determine the type of a vector
x <- "letter"

typeof(x)
length(x)

y <- c(1:10)

typeof(y)
length(y)

```

A quick individual exercise: determine the type and length of a built-in variable "letters"

````{r}
letters # Letter is a built-in variable
typeof(letters)
length(letters)


```

#Important Types of Atomic Vector

1. Logical type

````{r}
# Simple definition
(x <- c(TRUE, FALSE, TRUE))
typeof(x)

# Logical operation

(y <- c(1:10) %% 3 == 0) # Notice that you can do mathematical operations directly using a vector

typeof(y)

```

A quick individual exercise: for a vector c(1:10), examine each value to see whether the value is an even number (TRUE) of not (FLASE), return a logical vector that informs the result. 

````{r}
(x <- c(1:10) %% 2 == 0)

```

1. Numeric type

````{r}
(x <- c(1))

typeof(x)

# See the power of "L"
(y <- c(1L,2L,3L))

typeof(y)

```

A quick individual exercise: define x to be sqrt(2)^2; define y to be an integer 2.
Compare to see if they precisely equal to each other? Do you understand why? 

````{r}

x <- sqrt(2)^2

y <- 2L

x == y

x - y
```


3. Character (most complex type of atomic vector because string can contain an arbitary amount of data)

````{r}

x <- "My name is John"

typeof(x)
length(x)

y <- c("My","name","is","John")

typeof(y)
length(y)
```

#Operation of Atomic Vectors

Converting vectors
as.logical()
as.integer()
as.double()
as.character()

````{r}
(x <- c(0,1,2))

typeof(x)

(y <- as.integer(x))

typeof(y)

(z <- as.character(x))

typeof(z)

(a <- as.logical(x))

typeof(a)

```

Individual exercise: for a vector c("x","y","z"), convert is to integer, double, and logical types and see what you get. 

````{r}
(x <- c("x","y","z"))

typeof(x)
(y <- as.integer(x))
typeof(y)
(z <- as.double(x))
typeof(z)
(a <- as.logical(x))
typeof(a)

```

Testing vectors
is.logical()
is.integer()
is.double()
is.character()

````{r}
(x <- c(0,1,2))

is.integer(x)

is.double(x)

is.character(x)

is.logical(x)

```

Naming vectors
set_names
````{r}

(x <- c(Mon = 65, Tues = 62, Wed = 70))

(y <- set_names(c(65,62,70), c("Mon", "Tues", "Wed")))
```

Individual exercise: check the type of y. Can you see what's the benefit of defining a name to a variable?

````{r}

(y <- set_names(c(65,62,70), c("Mon", "Tues", "Wed")))
typeof(y)

y["Mon"]
y[2]
y[1]

```
Subsetting vectors
Using square bracket []

````{r}
x <- c("one", "two", "three", "four", "five")
x[c(3, 4, 5)]
x[c(-1,-2)]
#x[c(1,2,-3)]

(y <- set_names(c(65,62,70), c("Mon", "Tues", "Wed")))
y[c("Mon", "Tues")]

```
Individual exercise: see the power of logical operator in subset! 

The vector, x, is a random set of integers within 100. Can you find all the even values in the vector x? 

Hint-1: numeric vectors can undergo mathematical operations.
Hint-2: %% operator can help you get the remainder.

````{r}
x <- c(sample(100))
x[x %% 2 == 0]

```







