---
title: "Transform"
author: "John Yang"
output: html_notebook
---

# First install the packages
```{r}
install.packages("tidyverse")
install.packages("nycflights13")
```

# Now call the libraries

````{r}
library(nycflights13)
library(tidyverse)
```

# Now define a tibble. Pay attention to the differnt types of variables [integers<int>, doubles<dbl>, characters<chr>, data-times<dttm>, logical<lgl>, factors<fctr>, dates<date>]. 

````{r}
flights <- flights
flights 
# State the obvious~
# How many variables do we have in the tibble?
# How many observations do we have in the tibble?
```

# Let's try out filter() first

For the flights that got delayed, were they more likely to make up or get further delayed?
Hint-1: how to quantitatively define "delay"?
Hint-2: how to define "make up" and "further delayed"?

````{r}
( further_delay_0 = filter(flights, dep_delay > 0 & arr_delay-dep_delay > 0))
( make_up_0 = filter(flights, dep_delay > 0 & arr_delay-dep_delay < 0))
```


What are potential concerns in this analysis?


For example, what about making up or getting further delayed by 30 mins?

````{r}
( further_delay_30 = filter(flights, dep_delay > 0 & arr_delay-dep_delay >= 30))
( make_up_30 = filter(flights, dep_delay > 0 & arr_delay-dep_delay <= -30))
```

Another example, what about making up or getting further delayed by 50 mins?

````{r}
( further_delay_50 = filter(flights, dep_delay >= 0 & arr_delay-dep_delay >= 50))
( make_up_50 = filter(flights, dep_delay >= 0 & arr_delay-dep_delay <= -50))
```

# Now, let's work on arrange()

````{r}
# Sort flights to find the shortest flight.
arrange(flights, air_time)
# Sort flights to find the most delayed departure flight.
arrange(flights, desc(dep_delay))
```

Individual exercise (3 mins) 

Sort flights to find the fastest (highest speed) flight in July.
hint: 1. Speed is not a variable in the existing data table.
hint: 2. Speed is defined as "distance" over "air_time". Both variables can be found in flights.

````{r}
july1 = filter(flights, month == 7)
arrange(july1, desc(distance/air_time) )
```

# Find the variables you are interested in using select(). Note: Be aware of the number of variables.

````{r}
flights
# select three variables year, month, and day
select(flights, year, month, day)
# you can use "-" to eliminate variables
select(flights, -(year:dep_time))
# helper functions: contains, starts_with, ends_with
select(flights, contains("dep"))
    # select(flights, starts_with("dep"))
    # select(flights, ends_with("dep"))
# define a vector first and then use helper function any_of()
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, any_of(vars))
```

# Add new variables with mutate(). Add new variables that are derived from existing variables. 

````{r}
# Here we want to create a new table. 
# The new table, flights_speed, will contain multiple variables from flights.
# The new table will also add one more variable "speed" that is derived from distance and air_time
flights_speed <- select(flights, year:day, distance, air_time)
mutate(flights_speed, speed_mile_per_mins  = distance / air_time, speed_mile_per_hours = distance / air_time * 60)
```

Group exercise (7 mins):

1. Create a new tibble, flights_ind, to include year, month, day, and the variables that end with "delay".
2. Find all flights that departed in summer, and save your results to a new tibble, flights_ind_summer
3. Add a new variable to flights_ind_summer, time_gain, that represents the time gain during the flight.
4. Arrange the table based on the time_gain and put the 

````{r}
(flights_ind <-select(flights, year, month, day, ends_with("delay")))
(flights_ind_summer <- filter(flights_ind, month == 7 | month == 8 | month == 9 ))
(flights_ind_summer <- mutate(flights_ind_summer, time_gain = dep_delay - arr_delay))

```

# Finally, Group_by and summarise()

````{r}
# Get statistical information using summarise() 
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```
````{r}
# Get statistical information for a certain group by combining summarise() with group_by()
by_day = group_by(flights, year, month)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

````{r}
# Another example
by_origin = group_by(flights, origin)
summarise(by_origin, delay = mean(dep_delay, na.rm = TRUE))
```

Question: how many groups do you expect to see in the following command? 12+3 =15 or 12*3 = 36

````{r}
by_origin_month = group_by(flights, year, month, origin)
summarise(by_origin_month, delay = mean(dep_delay, na.rm = TRUE))
```

Group Challenge: Explore the relationship between the distance and average delay for each location in July？
Hint: Think about the order of verbs (summarise, group_by, filter, etc.) you will use to transform data.

````{r}

```
