---
title: "Transform-Assert"
author: "John Yang"
output: html_notebook
---

# First install the packages
```{r}
install.packages("tidyverse")
install.packages("nycflights13")
```
# Now call the libraries

````{r}
library(nycflights13)
library(tidyverse)
library(tidyr)
```

# Functional programming: Conditional Execution

Four variations:

1. if () {}

Example-1: Just "if"

````{r}
x <- "True"

if ( x == "True") {
  print("Yes!")
} 

```

Example-2: Multiple "if" statements

````{r}
x <- "True"

if ( x == "True") {
  print("Yes!")
} 
if ( x != 0) {
  print("Yes!")
}
if ( is.character(x)) {
  print("No!")
}
```

2. If () {} else if () {}

````{r}
x <- "True"
# x <- "Maybe"
# x <- "False"

if ( x == "True") {
  print("Yes!")
} else if ( x != "False"){
  print("Not false!")
}

```

3. if () {} else {}

````{r}
x <- "True"

if ( x == "True") {
  print("Yes!")
} else {
  print("No!")
}

```

4. If () {} else if () {} else {}

````{r}
x <- "True"

if ( x == "True") {
  print("Yes!")
} else if (x == "Maybe") {
  print("Maybe!")
} else {
  print("No!")
}

```

Individual excercise: Write a function named "judgement" to take three numeric arguments, a, b, and c. Determine whether a+b == c? Print "Yes!" or "No!", based on the judgement result. 

Hint: template of function definition: 
      function_name <- function (function_argument) {
              function_body
                                       }

````{r}
judgement <- function(a, b, c) {

}

judgement(1,2,3)
```

# Iteration

Motivation

````{r}
df <- list(
  a = 1:10,
  b = 11:20,
  c = 21:30,
  d = 31:40
)

df
```

We defined a function to scale atomic vectors:

````{r}
rescale_10 <- function(x) {
  x*10
}

```

Simplest example:

````{r}
nums = c()
for ( i in 1:3) {
  nums[i] <- i
}
nums
```

Looping a list:

````{r}
fruits <-list("apple","banana","cherry")

for (x in fruits) {
  print(x)
  }
```

Looping a list:

````{r}
fruits <-list("apple","banana","cherry")

for (x in seq_along(fruits)) {
  print(fruits[x])
  }
```

Individual exercise: examine the data types in flights!

Use For Loop to print the type of each column in nycflights13::flights

Hints: 
1. "i in flights" will iterate ever column in flights 
2. typeof() function can return the type of a variable

````{r}
#flights



```

Integrating condition and iteration to create more powerful code!

rescale_10: we will prepare a new rescale_10 function that can take either a numeric type or a list type, rescale the values stored inside by 10, and print the outcome.

````{r}
df <- list(
  a = 1:10,
  b = 11:20,
  c = 21:30,
  d = 31:40
)

rescale_10 <- function(x) {
  if (is.numeric(x)) {
    x = x*10
  } else if (is.list(x)) {
    for ( i in seq_along(x)) {
      x[[i]] = x[[i]]*10
      } 
    } else {
      print("Something is wrong. The variable can not be scaled!")
    }
  
  print(x)
}
# case-1: just an atomic vector
rescale_10(c(1,2,3))
# case-2: give the function a list
rescale_10(df)
# case-3: give the function a character
rescale_10("wonderful")
```

# Return value

````{r}
rescale_10 <- function(x) {
  for ( i in seq_along(x)) {
      x[[i]] = x[[i]]*10
  }
}

rescale_10_return <- function(x) {
  for ( i in seq_along(x)) {
      x[[i]] = x[[i]]*10
  }
  return(x)
}

#A <- rescale_10(list(1,2,3))
#print(A)

B <- rescale_10_return(list(1,2,3))
print(B)
```

Individual exercise: revise the following code to return the list after rescaling. 

````{r}
df <- list(
  a = 1:10,
  b = 11:20,
  c = 21:30,
  d = 31:40
)

rescale_10 <- function(x) {
  if (is.numeric(x)) {
    x = x*10
  } else if (is.list(x)) {
    for ( i in seq_along(x)) {
      x[[i]] = x[[i]]*10
      } 
    } else {
      print("Something is wrong. The variable can not be scaled!")
    }
  
}

C = rescale_10(df)
print(C)
```