---
title: "play_with_dataframe"
output: html_notebook
---

# install the packages
```{r}
install.packages("tidyverse")
install.packages("nycflights13")
```

# Now call the libraries
````{r}
library(nycflights13)
library(tidyverse)
```

````{r}
flights <- flights
flights 
```

# Let's try out filter() first

List all flights that departed on Jan 1, 2013

````{r}
( jan1 = filter(flights, month == 1, day == 1) )
```


A quick test:

Suppose we want to collect the flight information on the first two days of Jan (i.e. Jan 1st and Jan 2nd).
Which one of the following two commands will give us what we want?

Choice-A

````{r}
( jan1 = filter(flights, month == 1, day == 1 | day == 2) )
```

Choice-B

````{r}
( jan1 = filter(flights, month == 1, day == 1 & day == 2) )
```

2 mins individual exercise:

We want to find flights that aren't delayed on departure by more than two hours?

Hint-1: the tibble has a variable called dep_delay
Hint-2: the unit of the time variable is minute for the tibble. 

````{r}

```

Finally, if we have missing data, is.na() function will help us identify them. Like the following example:

````{r}

```

A quick recap for filter():

1. "==" is different from "="


2. Comparison operators: And Or Not


3. Be careful with natural vs. programming language.


4. is.na() will help us identify missing data.


Now, let's join the breakout room and solve problems with your teammate. 

Three filter() challenges:

1. Find the flights that flew from New York JFK to Houston (IAH or HOU)
2. Find the flights that departed in summer (July, August, and September)
3*. For the flights that got delayed, were they more likely to make up or get further delayed?
Hint-1: how to quantitatively define "delay"?
Hint-2: how to define "make up" and "further delayed"?

Flew from New York JFK to Houston (IAH or HOU)

````{r}

```

Departed in summer (July, August, and September)

````{r}

```

For the flights that got delayed, were they more likely to make up or get further delayed?
Hint-1: how to quantitatively define "delay"?
Hint-2: how to define "make up" and "further delayed"?

````{r}

```

What are potential concerns in this analysis?

# Now, let's work on arrange()

````{r}
arrange(flights, dep_delay)
```

Sort flights to find the fastest (highest speed) flights.

````{r}
arrange(flights, desc(distance/air_time))
```

Additional challenge:

Find the most delayed (departure) flight in July?
````{r}

```
