---
title: "Transform-Assert"
author: "John Yang"
output: html_notebook
---

# First install the packages
```{r}
install.packages("tidyverse")
install.packages("nycflights13")
```
# Now call the libraries

````{r}
library(nycflights13)
library(tidyverse)
library(tidyr)
```

# More excercises: Join and Transform!

````{r}
flights 
airlines 
airports 
planes 
weather
```
# Project: why flights delay?
# Aim-1: Maybe it's the weather condition to blame!


Task-1: Get the data tibble ready (5 mins)!
instructions: 
1. select the following variables "year:day", "hour", "origin", "dest", "tailnum", "carrier", "dep_delay", "arr_delay" from the tibble flights. The new tibble should be named "flights_sele"

2. left_join with weather tibble by "year", "month", "day", "hour", "origin". The new tibble will be named "flights_weather"

3. Check to make sure you have 20 variables in your final tibble.

````{r}
flights_sele <- flights %>% 
  select("year", "month", "day", "hour", "origin", "dest", "tailnum", "carrier", "dep_delay", "arr_delay")  
  
flights_weather <- flights_sele %>%  
  left_join(weather, by = c("year", "month", "day", "hour", "origin"))

flights_weather

```
Task-2: Analyze how certain variables of weather may influence the departure delay? Let's start from humidity! (5 mins)
instruction: 
1. Based on flights_weather, select "dep_delay" and "humid", and define a new tibble yes_delay that contains all cases with dep_delay > 0. Name it as yes_delay.

2. Based on flights_weather, select "dep_delay" and "humid", and define a new tibble yes_delay that contains all cases with dep_delay <= 0. Name it as no_delay.

3. Computing the mean and standard deviation of the humid for yes_delay and no_delay, remember to use na.rm=TRUE when performing the computation. 

4. Answer the question, does humidity infuences the departure delay?

````{r}
(yes_delay <- flights_weather %>% 
  select("dep_delay", "humid") %>% 
  mutate(delay_status = ifelse(dep_delay > 0, "Yes","No")) %>% 
  filter(delay_status == "Yes"))
  
(no_delay <-flights_weather %>% 
  select("dep_delay", "humid") %>% 
  mutate(delay_status = ifelse(dep_delay >= 0, "Yes","No")) %>% 
  filter(delay_status == "No"))

yes_delay %>% 
  summarise(mean_humid = mean(humid, na.rm = TRUE), sd = sd(humid, na.rm = TRUE))
no_delay %>% 
  summarise(mean_humid = mean(humid, na.rm = TRUE), sd = sd(humid, na.rm = TRUE))

```

Task-3: What about other factors, such as temperature and wind_speed? (5 mins)

````{r}
(yes_delay_temp_ws <- flights_weather %>% 
  select(dep_delay, temp, wind_speed) %>% 
  mutate(delay_status = ifelse(dep_delay > 0, "Yes","No")) %>% 
  filter(delay_status == "Yes"))
  
(no_delay_temp_ws <-flights_weather %>% 
  select(dep_delay, temp, wind_speed) %>% 
  mutate(delay_status = ifelse(dep_delay >= 0, "Yes","No")) %>% 
  filter(delay_status == "No"))

yes_delay_temp_ws %>% 
  summarise(mean_temp = mean(temp, na.rm = TRUE), sd_temp = sd(temp, na.rm = TRUE), mean_wind_speed = mean(wind_speed, na.rm = TRUE), sd_ws = sd(wind_speed, na.rm = TRUE))
no_delay_temp_ws %>% 
  summarise(mean_temp = mean(temp, na.rm = TRUE), sd_temp = sd(temp, na.rm = TRUE), mean_wind_speed = mean(wind_speed, na.rm = TRUE), sd_ws = sd(wind_speed, na.rm = TRUE))


```
```
# Project-2: How airport location affects the delay?

````{r}


```

