---
title: "Transform-Assert"
author: "John Yang"
output: html_notebook
---

# First install the packages
```{r}
install.packages("tidyverse")
install.packages("nycflights13")
```
# Now call the libraries

````{r}
library(nycflights13)
library(tidyverse)
library(tidyr)
```


# Lists

````{r}

# Remember if you want to define a vector of c(1,2,3)

(x <- c(1,2,3))

```
````{r}

# Remember if you want to define a vector of c(1,2,3)

(x <- list(1,2,3))

#x*10
```

````{r}
# Now let's see the power of list!
y <- list("a",1L,1.5, TRUE)
str(y)
```
````{r}
z <- list(list(1, 2), list(3, 4))
str(z)
```

Subseting elements in a list

````{r}
x <- list(1:3, "a string", pi, list(-1, -5))
str(x)
```

How to get the first element from the list?

First way:
````{r}
x[1]
```

Second way:
````{r}
x[[1]]
```

Third way:
````{r}
x <- list(a = 1:3, b = "a string", c = pi, d = list(-1, -5))

x$a
#x$1
```

# Function: Why we need that?

1. You can give a function a name that makes your code easier to understand

2. As requirements change, you only need to update the code in one place

3. You eliminate the chance of making incidental mistakes when you copy and paste the code.


````{r}
df <- list(
  a = 1:10,
  b = 11:20,
  c = 21:30,
  d = 31:40
)

df
```
Now we want to scale the first atomic vector of the list df$a by 10 times!

````{r}

(df$a <- df$a * 10)

```

Now we want to do the same thing in df$b df$c and df$d

But do you see what's wrong with the code?

````{r}

(df$b <- df$b * 10)
(df$c <- df$c * 10)
(df$d <- df$d * 10)

```
````{r}
df
```
It's time for a function!
#      function_name <- function (function_argument) {
#              function_body
#                                       }

Three important elements:
# a name
# argument
# function body

Here is a function we just defined:
````{r}
rescale_10 <- function(x) {
  x*10
}

test <- c(1,2,3)

rescale_10(test)
```

Now, with the function in hand, we are ready to apply the function to simplify the code structure!
````{r}
rescale_10(df$a)
rescale_10(df$b)
rescale_10(df$c)
rescale_10(df$d)
```

Individual excercise: Define a function named "rescale_01" to operate on a numeric type atomic vector. The function should scale every value in the vector by a factor of 0.1

````{r}
df <- list(
  a = 1:10,
  b = 11:20,
  c = 21:30,
  d = 31:40
)


# Your function goes here.



df
```
Some special aspects of function arguments in R

Let's write a function to compute x to the power of y. 

x^y 

````{r}

2^3

```

````{r}
pow <- function(x, y) {
result <- x^y
print(paste(x," raised to the power of ", y, " is ", result))
}

pow(2,3)
pow(x = 2, y = 3)
```

Now question to you (a poll!): 8 or 9?

We know that 2^3 = 8 and 3^2 = 9

First one: 8 or 9?
````{r}
pow(3, 2)
```

Second one: 8 or 9?
````{r}
pow(y = 3, x = 2)
```

Third one: 8 or 9?
````{r}
pow(3, x = 2)
```

Key: R doesn't care about the order of the arguments! It cares about the name of the variable! 

# Define the function to make it readable.

````{r}
# Too short
f()

# Not a verb, or descriptive
my_awesome_function()

# Long, but clear
classmate_names()
collapse_years()
```

# Functional programming: applying and choosing functions

Conditions that make your function smarter!

if...else structure

````{r}
x <- 1
# x <- 2
if ( x == 1) {
  print("Yes!")
} else {
  print("No!")
}

```

if...elif...else structure

````{r}
x <- 1
# x <- 2
# x <- 3
if ( x == 1) {
  print("Yes!")
} else if (x == 2) {
  print("Maybe!")
} else {
  print("No!")
}

```

Individual excercise: Write a function named "judgement" to take three numeric arguments, a, b, and c. Determine whether a+b == c? Print "yes" or "no", based on the judgement result. 

````{r}
# Your function goes here.


```