---
title: "Transform-Assert"
author: "John Yang"
output: html_notebook
---

# First install the packages
```{r}
install.packages("tidyverse")
install.packages("nycflights13")
```
# Now call the libraries

````{r}
library(nycflights13)
library(tidyverse)
library(tidyr)
```

# Defining four tibbles

````{r}
table1
table2
table3
table4a  # cases
table4b  # population
```

# First, let's do separation!

````{r}
table3
table3 %>% 
  separate(rate, into = c("cases", "population"), sep = "/", convert = TRUE)
```

Experimenting (1 min): what if you remove "convert = TRUE"?

````{r}
table3
table3 %>% 
  separate(rate, into = c("cases", "population"), sep = "/")
```

# Second, let's do unite!

````{r}
table3
table3 %>% 
  unite("country-year", country, year, sep = "-")
```

Individual Challenge (5 mins): Here we have a real life dataset about statistics of injury. 
First, select the "age_group" variable; 
Second, separate the age_group to two columns: "start_age" and "end_age"
Third, unite "start_age" and "end_age" back to one column named "age_group" using "/" as a separator.

````{r}
tbi_age <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-24/tbi_age.csv')
tbi_age
tbi_age %>% 

# Your code goes here:
select(age_group) %>% 
  separate(age_group, into = c("start_age", "end_age"), sep = "-", convert = TRUE) %>% 
  unite("age_group", start_age, end_age, sep = "/")
```

# Third, let's do pivot!

# pivot_longer: converting NAMEs to values

````{r}
table4a
table4a %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")
```

Individual challenge (2 mins): applying pivot_longer to convert the names of table4b to "year" and the existing values to "population"

````{r}
table4b 

# Your code goes here:
table4b %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "population")
```

# pivot_longer: converting NAMEs from values

````{r}
table2
table2 %>%
  
    pivot_wider(names_from = type, values_from = count)
```

Individual challenge (wider or longer?): For the following tibble, `1` and `2` refer to the first and second half of a year, respectively. The values under `1` and `2` refers to the rate of return. Please tidy the data by redefining the names to be "half" and "return".

````{r}
stocks <- tibble(
  year  = c(2015, 2016),
  `1` = c(1.88, 0.92),
  `2` = c(0.59, 0.17)
)
stocks

# Your code goes here:
stocks %>% 
  pivot_longer(`1`:`2`, names_to = "half", values_to = "return")

```

assert_rows(col_concat, is_uniq, Plant,Type,Treatment,conc) 
assert(not_na, Plant)

