---
title: "Transform-Assert"
author: "John Yang"
output: html_notebook
---

# First install the packages
```{r}
install.packages("tidyverse")
install.packages("nycflights13")
install.packages("covid19.analytics")
```

````{r}
library(covid19.analytics)
library(tidyverse)
```

Group challenge*: Identify the top ten states that have the highest mean growth rate of COVID cases in August, 2021.  
Hint-1: across() allows arithmetic operations for a range of variables. 
Hint-2: rowMeans() will return the by-row average of values in a tibble

````{r}
CV_US %>% 
   group_by(Province_State) %>% 
   summarise((across(`2021-08-01`:`2021-08-31`, sum) - across(`2021-07-31`:`2021-08-30`, sum))/across(`2021-07-31`:`2021-08-30`, sum)) %>%
   group_by(Province_State) %>%
   summarise( Mean_Aug_2021 = rowMeans(across(`2021-08-01`:`2021-08-31`), na.rm = TRUE)) %>%
   arrange(desc(Mean_Aug_2021))
```

````{r}
install.packages("assertr")
```

````{r}
library(assertr)
library(magrittr)
library(dplyr)
```

#Four functions assertr
verify: takes a data frame and a logical operation expression. 

# assert
# assert_rows
assert: takes a data frame and a "static" predicate.
predicate functions: not_na, within_bounds, in_set, is_uniq

# insist
# insist_rows
insist: takes a data frame and a "dynamic" predicate.
predicate functions: within_n_sds, within_n_mads


````{r}
mtcars
mtcars %>%
  verify(has_all_names("mpg", "vs", "am", "wt")) %>%
#  verify(has_all_names("mpg_error", "vs", "am", "wt")) %>%
  verify(nrow(.) > 10) %>%
# verify(nrow(.) > 10) %>%
# verify(mpg < 20) %>% # think about comparison to filter()
  verify(mpg > 0) 
```

Description of variables:

[1]	mpg	Miles/(US) gallon
[2]	cyl	Number of cylinders
[3]	disp	Displacement (cu.in.)
[4]	hp	Gross horsepower
[5]	drat	Rear axle ratio
[6]	wt	Weight (1000 lbs)
[7]	qsec	1/4 mile time
[8]	vs	Engine (0 = V-shaped, 1 = straight)
[9]	am	Transmission (0 = automatic, 1 = manual)
[10]	gear	Number of forward gears
[11]	carb	Number of carburetors

Individual excercise (5 mins): 
1. verify if the engine is either V-shaped or straight?
Hint: the variable, vs,  equals to 0 or 1.
2. also try verifying if the number of forward gears is between 3 and 5.

````{r}

mtcars %>%
  verify(vs == 0|vs==1) %>% 
  verify(gear>=3 & gear<=5)
 

```

Now let's take a look at assert

````{r}
mtcars %>%
  assert(in_set(0,1), am, vs) %>%
  #assert(in_set(0,1), cyl) %>% 
  #assert(within_bounds(9,10), gear, carb)
  assert(within_bounds(0,10), gear, carb)
```

Individual excercise (2 mins): 
assert to see if each car has a unique 1/4 mile time
Hint: use is_uniq()

````{r}
mtcars %>%
  assert(is_uniq, qsec) 
```

Now let's take a look at insist

````{r}
mtcars %>%
  insist(within_n_sds(3), mpg) %>%
  insist(within_n_mads(3), mpg)
```

Also - we have row operations!

````{r}
mtcars %>%
  #assert_rows(col_concat, is_uniq, vs, am) %>%
  assert_rows(col_concat, is_uniq, mpg, am, wt)
```

Finally, you can use chain_start and chain_end to examine all conditions at once. 

````{r}
mtcars %>%
       chain_start %>%
       verify(nrow(.) >= 10) %>%
       verify(mpg < 32) %>%
       assert(in_set(1, 2, 3, 4), carb) %>%
       assert_rows(col_concat, is_uniq, vs, am) %>%
       chain_end
```



